Prototyping Collision-Free MAC Protocols in Real Hardware

### 2013-10-18 journal.tex
Rearranged the introduction

\boldmath Collisions are a main cause of throughput degradation in WLANs. The current contention mechanism used in this type of network called Carrier Sense Multiple Access with Collision Avoidance (CSMA/CA) uses a Binary Exponential Backoff (BEB) mechanism to delay each contender attempt of transmitting, effectively reducing the collision probability. Nevertheless, CSMA/CA relies on a random backoff which in principle is unable to eliminate collisions, resulting in a network throughput degradation as more contenders attempt to share the channel. Carrier Sense Multiple Access with Enhanced Collision Avoindance (CSMA/ECA) is able to create a collision-free schedule in a totally distributed manner by means of picking a deterministic backoff after successfull transmissions. CSMA/ECA is able to support many contenders in a collision-free schedule, surpassing the achieved throughput of CSMA/CA and provide short-term throughput fairness among contenders.

This work reviews CSMA/ECA mechanisms and provides insightful simulations and the first real-life tests results that reveal its benefits over CSMA/CA under different network traffic conditions.

@@

\section{Introduction}\label{introduction}
\section{Related Work}\label{relatedWork}
\section{Simulation Results}\label{simulations}
\section{Prototyping CSMA/ECA}\label{prototype}
\section{CSMA/ECA in real hardware results}\label{prototypeResults}
\section{Conclusions}\label{conclusions}
\section{Acknowledgements}

>>>>

\boldmath Collisions are a main cause of throughput degradation in WLANs. The current contention mechanism used in this type of network called Carrier Sense Multiple Access with Collision Avoidance (CSMA/CA) uses a Binary Exponential Backoff (BEB) technique to delay each contender attempt of transmitting, effectively reducing the collision probability. Nevertheless, CSMA/CA relies on a random backoff that while effective and totally distributed, in principle is unable to eliminate collisions; degrading the network throughput as more contenders attempt to share the channel. Carrier Sense Multiple Access with Enhanced Collision Avoindance (CSMA/ECA) is able to create a collision-free schedule in a totally distributed manner by means of picking a deterministic backoff after successful transmissions. CSMA/ECA is able to support many contenders in a collision-free schedule, surpassing the achieved throughput of CSMA/CA and provides short-term throughput fairness among contenders.

This work reviews CSMA/ECA, providing insightful simulation results revealing its advantages over CSMA/CA. It also shows the first real-hardware implementation tests results of CSMA/ECA under saturated and unsaturated conditions.

@@

\section{Introduction}\label{introduction}
Wireless Local Area Networks (WLANs or IEEE 802.11 networks~\cite{802Standards}) are a popular solution for wireless connectivity, whereas in public places, work environments or at home. This technology works over an unlicensed spectrum in the Industrial, Scientific and Medical (ISM) radio bands (at around $2.4$ or $5$~GHz), which is a main reason for its popularity.

The Medium Access Control (MAC) scheme used in WLANs is called Distributed Coordination Function (DCF) and is based on Carrier Sense Multiple Access with Collision Avoidance (CSMA/CA) protocol. It has been widely adopted by manufacturers and consumers, making it very cheap to implement and an ubiquitous technology. Nevertheless, ever-growing throughput demands from upper layers have proven to be limited by WLANs' MAC, which by its nature is prone to collisions that degrade the overall performance as more nodes join the network.

The research community has pushed forward many alternatives to the current MAC in WLANs~\cite{bharghavan1994map,wang2004ncr,cali2000dti,lopez-toledo2006aoi,
barcelo2008lba,bellalta2009vtc,HE,CSMA_ECA,L_MAC2,hui2011epp,barcelo2011tcf}, but when a proposal deviates too much from CSMA/CA or time-critical operations are modified, its hardware implementation as part of WLANs MAC often becomes unlikely~\cite{WMP}; the standardization process taking many years without certainty of approval~\cite{perahia2008ieee}.

%Backwards compatibility is paramount, mostly because the existing user base is very big. Also, a suitable candidate to replace CSMA/CA should increase the number of supported contenders while augmenting the offered throughput.

\section{Related Work}\label{relatedWork}
\section{Simulation Results}\label{simulations}
\section{Prototyping CSMA/ECA}\label{prototype}
\section{CSMA/ECA in real hardware results}\label{prototypeResults}
\section{Conclusions}\label{conclusions}
\section{Acknowledgements}
